---
title: "Stat 695 Project Markdown File"
author: "Dylan Clarke"
date: "2025-05-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

We begin by loading in the data and packages we need for this project.

```{r preliminaries, echo=FALSE}
library(causaldata)
library(survival)
library(ggplot2)
library(tidyverse)
library("survminer")
library("splitstackshape")
library(olsrr)
library(ipw)
library(timereg)
library(MatchIt)
library(xtable)
library(rifttable)
library(confintr)

setwd("~/R/win-library/4.0/Causal Survival")

nhefs = causaldata::nhefs_complete
nhefs = subset(nhefs,!is.na(nhefs$income))
```

Now we clean the data a bit to get it ready. This creates the survival variable we will use.

```{r cleaning}
start_date = ISOdate(1983,1,1)
nhefs$death_date = ISOdate(paste0(19,nhefs$yrdth),nhefs$modth,nhefs$dadth)
len_study = as.numeric(ISOdate(1993,1,1) - start_date)
nhefs$surv = as.numeric(nhefs$death_date - start_date)
nhefs$surv[is.na(nhefs$surv)] = len_study
nhefs$survival = with(nhefs,Surv(surv,death))

nhefs$survtime <- ifelse(nhefs$death==0, 120, 
                         (nhefs$yrdth-83)*12+nhefs$modth)
```

## Inital Exploration of the Data

## Correlations

```{r corr with qsmk}
cor(Filter(is.numeric,nhefs))

cor(as.numeric(nhefs$sex),nhefs$qsmk)
cor(as.numeric(nhefs$race),nhefs$qsmk)
cor(as.numeric(nhefs$active),nhefs$qsmk)
cor(as.numeric(nhefs$active),nhefs$death)
cor(as.numeric(nhefs$exercise),nhefs$qsmk)
cor(as.numeric(nhefs$exercise),nhefs$death)
```
### Selecting Covariates

```{r select covariates}
sapply(nhefs[,24:57],try(table,silent = TRUE))

qsmk_mod = glm(qsmk~ sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + price82 + tax82 + pepticulcer + alcoholpy + otherpain  + exercise + marital,
               data = nhefs, family = binomial)

death_mod = glm(death~sex + age + race + income + school + ht + wt82 + smokeintensity + smokeyrs + price82 + tax82 + bronch + pepticulcer + hayfever + alcoholpy + alcoholfreq + alcoholtype + alcoholhowmuch + headache + otherpain + nerves + infection + active  + exercise + cholesterol,
               data = nhefs, family = binomial)

ols_step_backward_p(qsmk_mod)
```


##Inverse Probability Weights (IPTW, IPCW)

```{r IPW}
#DOES NOT RUN
#IPTW
ipw_treat = ipwpoint(
  exposure   = qsmk,
  family     = "binomial",
  link       = "logit",
  denominator = ~ sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
  numerator   = ~ 1,
  data       = nhefs
)

qsmk_mod = glm(qsmk~ sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
               data = nhefs, family = binomial)

nhefs$p = fitted(qsmk_mod)

pqsmk = mean(nhefs$qsmk)  
nhefs$iptw <- ifelse(nhefs$qsmk==1,  pqsmk/nhefs$p,  
                                (1-pqsmk)/(1-nhefs$p))

#IPCW
nhefs$censored = 1 - nhefs$death
mod_cens = coxph(
  Surv(surv, death, type = "right") ~ qsmk + sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
  data = nhefs
)

bh = basehaz(mod_cens, centered = FALSE)

lp = predict(mod_cens, type = "lp")

H0_at_t <- approx(
  x = bh$time,
  y = bh$hazard,
  xout = nhefs$surv,
  method = "constant",
  rule = 2
)$y

nhefs$p_cens = exp(- H0_at_t * exp(lp))

pC = mean(nhefs$death == 0)
nhefs$ipcw = ifelse(
  nhefs$death == 0,
  pC / nhefs$p_cens,
  1 / (1 - nhefs$p_cens)
)

nhefs$w = nhefs$ipcw * nhefs$iptw

#Version 2
p.denom <- glm(qsmk ~ sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital, 
               data=nhefs, family=binomial())
nhefs$pd.qsmk <- predict(p.denom, nhefs, type="response")

p.num <- glm(qsmk ~ 1, data=nhefs, family=binomial())
nhefs$pn.qsmk <- predict(p.num, nhefs, type="response")

# computation of estimated weights
nhefs$sw.a2 <- ifelse(nhefs$qsmk==1, nhefs$pn.qsmk/nhefs$pd.qsmk,
                     (1-nhefs$pn.qsmk)/(1-nhefs$pd.qsmk))

# estimation of denominator of ip weights
p.denom <- glm(qsmk ~ age + sex + wt71 +income + smokeintensity + smokeyrs, 
               data=nhefs, family=binomial())
nhefs$pd.qsmk <- predict(p.denom, nhefs, type="response")

# estimation of numerator of ip weights
p.num <- glm(qsmk ~ 1, data=nhefs, family=binomial())
nhefs$pn.qsmk <- predict(p.num, nhefs, type="response")

# computation of estimated weights
nhefs$sw.a <- ifelse(nhefs$qsmk==1, nhefs$pn.qsmk/nhefs$pd.qsmk,
                     (1-nhefs$pn.qsmk)/(1-nhefs$pd.qsmk))
```

## Aalen Model Discarded Attempt

```{r aalen}
#Baseline Aalen
aalen_model_uw = aareg(survival~ qsmk + sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
      data = nhefs)

summary(aalen_model_uw)

times = aalen_model_uw$times

coef = aalen_model_uw$coefficient

aalen_model_uw

Lambda0 = coef[,"Intercept"]
Bqmk = coef[, "qsmk"]
BS = coef[,"sex1"]
BR = coef[,"race1"]
BA = coef[, "age"]
BSc = coef[,"school"]
BH = coef[,"ht"]
BW = coef[,"wt82"]
BWc = coef[,"wt82_71"]
BSi = coef[,"smokeintensity"]
BSy = coef[,"smokeyrs"]
BPu = coef[,"pepticulcer"]
BAy = coef[,"alcoholpy"]
BO = coef[,"otherpain"]
BE1 = coef[,"exercise1"]
BE2 = coef[,"exercise2"]
BM = coef[,"marital"]

n = nrow(nhefs)
S0_mat = matrix(0, nrow = n, ncol = length(times))
S1_mat = matrix(0, nrow = n, ncol = length(times))

for (j in seq_along(times)) {
  linpred0 =          BS[j]*(nhefs$sex == 1) +
    BR[j]*(nhefs$race == 1) +
    BA[j]*(nhefs$age == 1) +
    BSc[j]*nhefs$school +
    BH[j]*nhefs$ht +
    BW[j]*nhefs$wt82 +
    BWc[j]*nhefs$wt82_71 + 
    BSi[j]*nhefs$smokeintensity +
    BSy[j]*nhefs$smokeyrs +
    #BP[j]*nhefs$price82 +
    BPu[j]*nhefs$pepticulcer +
    BAy[j]*nhefs$alcoholpy +
    BO[j]*nhefs$otherpain +
    BE1[j]*(nhefs$exercise == 1) +
    BE2[j]*(nhefs$exercise == 2) +
    BM[j]*nhefs$marital
  linpred1 = Bqmk[j] +  BS[j]*(nhefs$sex == 1) +
    BR[j]*(nhefs$race == 1) +
    BA[j]*(nhefs$age == 1) +
    BSc[j]*nhefs$school +
    BH[j]*nhefs$ht +
    BW[j]*nhefs$wt82 +
    BWc[j]*nhefs$wt82_71 + 
    BSi[j]*nhefs$smokeintensity +
    BSy[j]*nhefs$smokeyrs +
    #BP[j]*nhefs$price82 +
    BPu[j]*nhefs$pepticulcer +
    BAy[j]*nhefs$alcoholpy +
    BO[j]*nhefs$otherpain +
    BE1[j]*(nhefs$exercise == 1) +
    BE2[j]*(nhefs$exercise == 2) +
    BM[j]*nhefs$marital
  
  S0_mat[, j] <- exp(Lambda0[j] + linpred0)
  S1_mat[, j] <- exp(Lambda0[j] + linpred1)
}

S0 <- colMeans(S0_mat,na.rm = TRUE)
S1 <- colMeans(S1_mat, na.rm = TRUE)

cumprod(S0)
cumprod(S1)


#IPW Aalen
aalen_model_w = aareg(survival~ qsmk,
      data = nhefs, weights = sw.a2)

times = aalen_model_w$times

coef = aalen_model_w$coefficient

Lambda0 = coef[,"Intercept"]
Bqmk = coef[, "qsmk"]

n = nrow(nhefs)
S0_mat = matrix(0, nrow = n, ncol = length(times))
S1_mat = matrix(0, nrow = n, ncol = length(times))

for (j in seq_along(times)) {
  linpred1 = Bqmk[j] 
  
  S0_mat[, j] <- exp(-Lambda0[j])
  S1_mat[, j] <- exp(-Lambda0[j] - linpred1)
}

S0 <- colMeans(S0_mat,na.rm = TRUE)
S1 <- colMeans(S1_mat, na.rm = TRUE)

cumprod(S0)[291]
cumprod(S1)[291]

surv_diff_aalen_w = cumprod(S1)[291] - cumprod(S0)[291]

#Bootstrap:
bootaalen <- function(){
  sam = sample(nhefs$seqn,length(unique(nhefs$seqn)),replace = TRUE)
   data = nhefs %>% filter(seqn %in% sam)
  data <- df_bs %>%
    group_by(seqn) %>%
    mutate(dup = sum(seqn == sam)) %>%
    ungroup() %>%
    uncount(weights = dup)
  p.denom <- glm(qsmk ~ sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital, 
               data=data, family=binomial())
data$pd.qsmk <- predict(p.denom, data, type="response")

p.num <- glm(qsmk ~ 1, data=data, family=binomial())
data$pn.qsmk <- predict(p.num, data, type="response")

data$sw.a2 <- ifelse(data$qsmk==1, data$pn.qsmk/data$pd.qsmk,
                     (1-data$pn.qsmk)/(1-data$pd.qsmk))

aalen_model_w = aareg(survival~ qsmk + sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
      data = data, weights = sw.a2)

times = aalen_model_w$times

coef = aalen_model_w$coefficient

Lambda0 = coef[,"Intercept"]
Bqmk = coef[, "qsmk"]
BS = coef[,"sex1"]
BR = coef[,"race1"]
BA = coef[, "age"]
BSc = coef[,"school"]
BH = coef[,"ht"]
BW = coef[,"wt82"]
BWc = coef[,"wt82_71"]
BSi = coef[,"smokeintensity"]
BSy = coef[,"smokeyrs"]
BPu = coef[,"pepticulcer"]
BAy = coef[,"alcoholpy"]
BO = coef[,"otherpain"]
BE1 = coef[,"exercise1"]
BE2 = coef[,"exercise2"]
BM = coef[,"marital"]

n = nrow(data)
S0_mat = matrix(0, nrow = n, ncol = length(times))
S1_mat = matrix(0, nrow = n, ncol = length(times))

for (j in seq_along(times)) {
  linpred0 =          BS[j]*(data$sex == 1) +
    BR[j]*(data$race == 1) +
    BA[j]*(data$age == 1) +
    BSc[j]*data$school +
    BH[j]*data$ht +
    BW[j]*data$wt82 +
    BWc[j]*data$wt82_71 + 
    BSi[j]*data$smokeintensity +
    BSy[j]*data$smokeyrs +
    #BP[j]*data$price82 +
    BPu[j]*data$pepticulcer +
    BAy[j]*data$alcoholpy +
    BO[j]*data$otherpain +
    BE1[j]*(data$exercise == 1) +
    BE2[j]*(data$exercise == 2) +
    BM[j]*data$marital
  linpred1 = Bqmk[j] +  BS[j]*(data$sex == 1) +
    BR[j]*(data$race == 1) +
    BA[j]*(data$age == 1) +
    BSc[j]*data$school +
    BH[j]*data$ht +
    BW[j]*data$wt82 +
    BWc[j]*data$wt82_71 + 
    BSi[j]*data$smokeintensity +
    BSy[j]*data$smokeyrs +
    #BP[j]*data$price82 +
    BPu[j]*data$pepticulcer +
    BAy[j]*data$alcoholpy +
    BO[j]*data$otherpain +
    BE1[j]*(data$exercise == 1) +
    BE2[j]*(data$exercise == 2) +
    BM[j]*data$marital
  
  S0_mat[, j] <- exp(Lambda0[j] + linpred0)
  S1_mat[, j] <- exp(Lambda0[j] + linpred1)
}

S0 <- colMeans(S0_mat,na.rm = TRUE)
S1 <- colMeans(S1_mat, na.rm = TRUE)

cumprod(S0)[291]
cumprod(S1)[291]

return(cumprod(S1)[291] - cumprod(S0)[291])
}

#Run bootstrap
aalen_vals = vector("list",500)
for(i in 1:500){
  aalen_vals[i] = bootaalen()
}

aalen_vals2 = vector("list",500)
for(i in 1:500){
  aalen_vals2[i] = bootaalen()
}
```

## Marginal Structural Models

```{r}
# creation of person-month data
nhefs.ipw = expandRows(nhefs, "survtime", drop=F) 
nhefs.ipw$time = sequence(rle(as.numeric(nhefs.ipw$seqn))$lengths)-1
nhefs.ipw$event = ifelse(nhefs.ipw$time==nhefs.ipw$survtime-1 & 
                            nhefs.ipw$death==1, 1, 0)
nhefs.ipw$timesq <- nhefs.ipw$time^2

# fit of weighted hazards model
ipw.model <- glm(event==0 ~ qsmk + I(qsmk*time) + I(qsmk*timesq) + 
                   time + timesq, family=binomial(), weight=w,
                 data=nhefs.ipw)
summary(ipw.model)

ipw.model_2 <- glm(event==0 ~ qsmk + I(qsmk*time) + I(qsmk*timesq) + 
                   time + timesq, family=binomial(), weight=sw.a,
                 data=nhefs.ipw)
summary(ipw.model)

# creation of survival curves
ipw.qsmk0 <- data.frame(cbind(seq(0, 119),0,(seq(0, 119))^2))
ipw.qsmk1 <- data.frame(cbind(seq(0, 119),1,(seq(0, 119))^2))

colnames(ipw.qsmk0) <- c("time", "qsmk", "timesq")
colnames(ipw.qsmk1) <- c("time", "qsmk", "timesq")

# assignment of estimated (1-hazard) to each person-month */
ipw.qsmk0$p.noevent0 <- predict(ipw.model, ipw.qsmk0, type="response")
ipw.qsmk1$p.noevent1 <- predict(ipw.model, ipw.qsmk1, type="response")

# computation of survival for each person-month
ipw.qsmk0$surv0 <- cumprod(ipw.qsmk0$p.noevent0)
ipw.qsmk1$surv1 <- cumprod(ipw.qsmk1$p.noevent1)

# some data management to plot estimated survival curves
ipw.graph <- merge(ipw.qsmk0, ipw.qsmk1, by=c("time", "timesq"))
ipw.graph$survdiff <- ipw.graph$surv1-ipw.graph$surv0

# plot
ggplot(ipw.graph, aes(x=time, y=surv)) + 
  geom_line(aes(y = surv0, colour = "0")) + 
  geom_line(aes(y = surv1, colour = "1")) + 
  xlab("Months") + 
  scale_x_continuous(limits = c(0, 120), breaks=seq(0,120,12)) +
  scale_y_continuous(limits=c(0.6, 1), breaks=seq(0.6, 1, 0.2)) +
  ylab("Survival") + 
  ggtitle("Survival from IP weighted hazards model") + 
  labs(colour="A:") +
  theme_bw() + 
  theme(legend.position="bottom")

S0 = ipw.qsmk0$surv0 
S1 = ipw.qsmk1$surv1

rmst0_msm = sum( (S0[-1] + S0[-length(S0)])/2 )
rmst1_msm = sum( (S1[-1] + S1[-length(S1)])/2 )

ate_rmst_msm = (rmst1_msm - rmst0_msm) * (3653/120)
```
## Aalen II
```{r}
aalen_model_uw_II = aalen(survival~ qsmk + const(sex) + const(age) + const(wt71) + const(smokeintensity) + const(smokeyrs) +const(income),
      data = nhefs)

aalen_model_uw_base = aalen(survival~ qsmk,
      data = nhefs)

aalen_model_uw_III = aalen(survival~ qsmk,
      data = nhefs, weights = nhefs$sw.a ,max.time = 3653,resample.iid = TRUE,n.sim = 500,id = nhefs$seqn)

aalen_list = as.data.frame(aalen_model_uw_II$cum)

aalen_list_covariates = aalen_model_uw_II$gamma

means_covariates = nhefs %>% summarize(sex = mean(as.numeric(sex)),
                                       age = mean(age),
                                       wt71 = mean(wt71),
                                       smokeintensity = mean(smokeintensity),
                                       smokeyrs = mean(smokeyrs),
                                       income = mean(income))

covariate_factor = sum(unlist((as.matrix(means_covariates)) * t(as.matrix(aalen_list_covariates)))[1,])

frame_aalen_uw = aalen_list %>% mutate(est_0 = exp(-`(Intercept)` - covariate_factor),est_1 = exp(-`(Intercept)` - qsmk - covariate_factor))

tail(frame_aalen_uw$est_1,1) - tail(frame_aalen_uw$est_0,1)

frame_pred_uw_62 = 

aalen_list_II = as.data.frame(aalen_model_uw_III$cum)

frame_aalen = aalen_list_II %>% mutate(est_0 = exp(-`(Intercept)`),est_1 = exp(-`(Intercept)` - qsmk))

tail(frame_aalen$est_1,1) - tail(frame_aalen$est_0,1)

aalen_list_base = as.data.frame(aalen_model_uw_base$cum)

frame_aalen_base = aalen_list_base %>% mutate(est_0 = exp(-`(Intercept)`),est_1 = exp(-`(Intercept)` - qsmk))

tail(frame_aalen_base$est_1,1) - tail(frame_aalen_base$est_0,1)

boot_aalen = aalen_model_uw_III$B.iid

predict.aalen(aalen_model_uw_III,nhefs)

aalen_vals = vector("list",500)
z = 1
for(j in boot_aalen){
  aalen_vals[z] = j[]
  z = z +1
}

ggplot(frame_aalen_base,aes(time)) + 
  geom_line(aes(y = est_0,colour = "est_0")) +
  geom_line(aes(y = est_1,colour = "est_1")) +
  labs(title = "Aalen Esimated Survival Curves for Unweighted Covariate Free Model") + 
  ylab("Survival Probability") +
  scale_color_hue(labels=c("est_0" = "qsmk = 0", "est_1"="qsmk = 1"))

#roughly estimated standard error
sqrt(sum(aalen_model_uw_base$var.cum[279,c(2,3)]))

#estimate
tail(frame_aalen_base$est_1,1) - tail(frame_aalen_base$est_0,1)

ggplot(frame_aalen,aes(time)) + 
  geom_line(aes(y = est_0,colour = "est_0")) +
  geom_line(aes(y = est_1,colour = "est_1")) +
  labs(title = "Aalen Esimated Survival Curves for Weighted Covariate Free Model") + 
  ylab("Survival Probability") +
  scale_color_hue(labels=c("est_0" = "qsmk = 0", "est_1"="qsmk = 1"))

#roughly estimated standard error
sqrt(sum(aalen_model_uw_III$var.cum[279,c(2,3)]))

#estimate
tail(frame_aalen$est_1,1) - tail(frame_aalen$est_0,1)

ggplot(frame_aalen_uw,aes(time)) + 
  geom_line(aes(y = est_0,colour = "est_0")) +
  geom_line(aes(y = est_1,colour = "est_1")) +
  labs(title = "Aalen Esimated Survival Curves for Unweighted Model with Covariates") + 
  ylab("Survival Probability") +
  scale_color_hue(labels=c("est_0" = "qsmk = 0", "est_1"="qsmk = 1"))

#roughly estimated standard error
sqrt(sum(c(as.matrix(means_covariates) %*% as.matrix(aalen_model_uw_II$var.gamma) %*% t(as.matrix(means_covariates)),aalen_model_uw_II$var.cum[279,c(2,3)])))

#estimate
tail(frame_aalen_uw$est_1,1) - tail(frame_aalen_uw$est_0,1)
```

## Cox Proportional Hazards

```{r}
cox_uw = coxph(survival ~ qsmk,data = nhefs, weights = sw.a2)

surv_adjustedcurves(cox_uw,variable = qsmk,data = nhefs)
```

## Plugin and Doubly Robust Estimator

```{r}
#We need to estimate the conditional probability of treatment assignment and survival for each subject to adminstrative censoring

#Creation of instrumental variable - change in price with tax #OLD
#nhefs$price_change = nhefs$price71_82 + nhefs$tax71_82
#nhefs$price_change[is.na(nhefs$price_change)] #<- mean(na.omit(nhefs$price_change)) #Set price change to 

nhefs$hightax82[is.na(nhefs$hightax82)]<- 0
cor(as.numeric(nhefs$hightax82),nhefs$qsmk)
cor(as.numeric(nhefs$hightax82),nhefs$death)

ci_cor(as.numeric(nhefs$hightax82),nhefs$qsmk)
ci_cor(as.numeric(nhefs$hightax82),nhefs$death)

#Probability of treatment assignment:
qsmk_mod = glm(qsmk~ hightax82 + age + sex + wt71 + income + smokeintensity + smokeyrs,
               data = nhefs, family = binomial)

#Create matrix for imputation

#old: imput_data = nhefs %>% select("sex","race","age","school","ht","wt82","wt82_71","smokeintensity","smokeyrs","pepticulcer","alcoholpy","otherpain","exercise","marital")
imput_data = nhefs %>% select("sex","age","wt71","smokeintensity","smokeyrs","income")

imput_data_1 = imput_data %>% mutate("hightax82" = 1)
imput_data_0 = imput_data %>% mutate("hightax82" = 0)

comp = data.frame(z1 = predict(qsmk_mod, newdata = imput_data_1,type = "response"),z0 = predict(qsmk_mod, newdata = imput_data_0,type = "response"))

#Probability of survival (or death) to end of study:
nhefs$live = 1 - nhefs$death
death_mod = glm(death~ hightax82 + age + sex + wt71 + income + smokeintensity + smokeyrs,
               data = nhefs, family = binomial)

imput_data_1_d = imput_data %>% mutate("hightax82" = 1)
imput_data_0_d = imput_data %>% mutate("hightax82" = 0)

comp_d = data.frame(d1 = predict(death_mod, newdata = imput_data_1_d,type = "response"),d0 = predict(death_mod, newdata = imput_data_0_d,type = "response"))

est_data = cbind(comp,comp_d)

est_data$est = (est_data$d1 - est_data$d0)/(est_data$z1 - est_data$z0)

#Plug-in Estimate:
mean((est_data$d1 - est_data$d0)/(est_data$z1 - est_data$z0))

#Doubly Robust was Abandoned- See paper
```

## Propensity Score Matching
```{r}
match_obj = matchit(qsmk~ hightax82 + sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
                    data = nhefs,method = "nearest", distance = "glm",
                    ratio = 1, replace = FALSE)

s_match = summary(match_obj)

s_match$sum.matched
print(xtable(s_match$sum.all, type = "latex"), file = "Initial_Cov_Balance_695.tex")
print(xtable(s_match$sum.matched, type = "latex"), file = "Matched_Cov_Balance_695.tex")

nhefs_match = match.data(match_obj)

#Version 2
match_obj_2 = matchit(qsmk~ age + wt82_71,
                    data = nhefs,method = "nearest", distance = "glm",
                    ratio = 1, replace = FALSE)

s_match_2 = summary(match_obj_2)

print(xtable(s_match$sum.all, type = "latex"), file = "Initial_Cov_Balance_695.tex")
print(xtable(s_match$sum.matched, type = "latex"), file = "Matched_Cov_Balance_695.tex")

nhefs_match_2 = match.data(match_obj_2)

#Version 3
match_obj_3 = matchit(qsmk~ + age + sex + wt71 + income + smokeintensity + smokeyrs,
                    data = nhefs,method = "nearest", distance = "glm",
                    ratio = 1, replace = FALSE, )

s_match_3 = summary(match_obj_3)

print(xtable(s_match_3$sum.all, type = "latex"), file = "Initial_Cov_Balance_695.tex")
print(xtable(s_match_3$sum.matched, type = "latex"), file = "Matched_Cov_Balance_695.tex")

nhefs_match_3 = match.data(match_obj_3)
```

## Analysis with Propensity Matching

```{r}
#Aalen on Matched
aalen_model_matched = aalen(survival~ qsmk,
      data = nhefs_match,resample.iid = TRUE,n.sim = 500,id = nhefs_match$seqn)

aalen_list_matched = as.data.frame(aalen_model_matched$cum)

frame_aalen_match = aalen_list_matched %>% mutate(est_0 = exp(-`(Intercept)`),est_1 = exp(-`(Intercept)` - qsmk))

tail(frame_aalen_match$est_1,1) - tail(frame_aalen_match$est_0,1)

ggplot(frame_aalen_match,aes(time)) + 
  geom_line(aes(y = est_0,colour = "est_0")) +
  geom_line(aes(y = est_1,colour = "est_1")) +
  labs(title = "Aalen Esimated Survival Curves for Propensity Matching") + 
  ylab("Survival Probability") +
  scale_color_hue(labels=c("est_0" = "qsmk = 0", "est_1"="qsmk = 1"))

#Aalen on Matched Version 2
aalen_model_matched_2 = aalen(survival~ qsmk,
      data = nhefs_match_2,resample.iid = TRUE,n.sim = 500,id = nhefs_match_2$seqn)

aalen_list_matched_2 = as.data.frame(aalen_model_matched_2$cum)

frame_aalen_match_2 = aalen_list_matched_2 %>% mutate(est_0 = exp(-`(Intercept)`),est_1 = exp(-`(Intercept)` - qsmk))

tail(frame_aalen_match_2$est_1,1) - tail(frame_aalen_match_2$est_0,1)

ggplot(frame_aalen_match_2,aes(time)) + 
  geom_line(aes(y = est_0,colour = "est_0")) +
  geom_line(aes(y = est_1,colour = "est_1")) +
  labs(title = "Aalen Esimated Survival Curves for Propensity Matching") + 
  ylab("Survival Probability") +
  scale_color_hue(labels=c("est_0" = "qsmk = 0", "est_1"="qsmk = 1"))

#Aalen on Matched Version 3
aalen_model_matched_3 = aalen(survival~ qsmk,
      data = nhefs_match_3,resample.iid = TRUE,n.sim = 500,id = nhefs_match_3$seqn)

aalen_list_matched_3 = as.data.frame(aalen_model_matched_3$cum)

frame_aalen_match_3 = aalen_list_matched_3 %>% mutate(est_0 = exp(-`(Intercept)`),est_1 = exp(-`(Intercept)` - qsmk))

tail(frame_aalen_match_3$est_1,1) - tail(frame_aalen_match_3$est_0,1)

ggplot(frame_aalen_match_3,aes(time)) + 
  geom_line(aes(y = est_0,colour = "est_0")) +
  geom_line(aes(y = est_1,colour = "est_1")) +
  labs(title = "Aalen Esimated Survival Curves for Propensity Matching") + 
  ylab("Survival Probability") +
  scale_color_hue(labels=c("est_0" = "qsmk = 0", "est_1"="qsmk = 1"))

#roughly estimated standard error
std_error_aalenweighted = sqrt(sum(aalen_model_matched_3$var.cum[-1,c(2,3)]))
sqrt(sum(aalen_model_matched_3$var.cum[159,c(2,3)]))

```

## Logistic with Maching

```{r}
#No matching
glm_match = glm(death~ qsmk + age + sex + wt71 + income + smokeintensity + smokeyrs,
               data = nhefs, family = binomial)

imput_data = nhefs %>% select("sex","age","wt71","smokeintensity","smokeyrs","income")

imput_data_1_q = imput_data %>% mutate("qsmk" = 1)
imput_data_0_q = imput_data %>% mutate("qsmk" = 0)

hat_s1 = predict(glm_match,imput_data_1_q,type = "response")
hat_s0 = predict(glm_match,imput_data_0_q,type = "response")

mean(hat_s1 - hat_s0)

#Matching
glm_matched = glm(death~ qsmk + age + sex + wt71 + income + smokeintensity + smokeyrs,
               data = nhefs_match_3, family = binomial)

hat_s12 = predict(glm_matched,imput_data_1_q,type = "response")
hat_s02 = predict(glm_matched,imput_data_0_q,type = "response")

mean(hat_s12 - hat_s02)
```

## Bootstrap

```{r}
bootlog <- function(){
  sam = sample(nhefs_match_3$seqn,length(unique(nhefs_match_3$seqn)),replace = TRUE)
   data = nhefs %>% filter(seqn %in% sam)
  data <- data %>%
    group_by(seqn) %>%
    mutate(dup = sum(seqn == sam)) %>%
    ungroup() %>%
    uncount(weights = dup)

glm_boot = glm(death~ qsmk + age + sex + wt71 + income + smokeintensity + smokeyrs,
               data = data, family = binomial)

hat_s12_boot = predict(glm_boot,imput_data_1_q,type = "response")
hat_s02_boot = predict(glm_boot,imput_data_0_q,type = "response")

return(mean(hat_s12_boot - hat_s02_boot))
}

boot_list = vector("list",500)
for(i in 1:500){
  boot_list[i] = bootlog()
}

sd(unlist(boot_list))
```


## Weighting Code

```{r}
library(tableone)
# Define variables you want in the table
vars <- c("wt71", "income", "sex", "age","smokeintensity","smokeyrs")

# Define the factor variable to stratify by (e.g., treatment group)


# Create Table 1 (basic)
table1 <- CreateTableOne(vars = vars, strata = "qsmk", data = nhefs)

# Print Table 1
print(table1, showAllLevels = TRUE,latex = T)
nhefs = nhefs[c("wt71","income","age","sex","smokeintensity","smokeyrs","qsmk","surv","death")]
nhefs = na.omit(nhefs)
ps_model <- glm(qsmk ~wt71 + income + sex + age + smokeintensity + smokeyrs, 
                data = nhefs, family = binomial(link = "logit") )
ps <- predict(ps_model, type = "response")
ipw <- ifelse(nhefs$qsmk == 1, 1/ps, 1/(1-ps))
p_treat = mean(nhefs$qsmk)
sw12 <- ifelse(nhefs$qsmk == 1, p_treat/ps, (1-p_treat)/(1-ps))
ow <- ifelse(nhefs$qsmk == 1, 1-ps,ps)

nhefs$ps = ps

ggplot(nhefs) + geom_histogram(aes(x = ps, col = as.factor(qsmk)))+facet_wrap(~ as.factor(qsmk), scales = "free") +
  theme_minimal()

km_ipw = survfit(Surv(as.numeric(surv),death) ~ qsmk, data = nhefs, weights = ipw)
km_ipw %>% ggsurvfit()

km_sw <- survfit(Surv(as.numeric(surv),death) ~ qsmk, data = nhefs, weights = sw12)
km_sw %>% ggsurvfit()

km_regular <-survfit(Surv(as.numeric(surv),death) ~ qsmk, data = nhefs)
km_regular %>% ggsurvfit()

km_sw$surv[c(193,279)]
km_sw$std.err[c(193,279)]
km_ipw$surv[c(193,279)]
km_ipw$std.err[c(193,279)]
## estimate 
0.82905 - 0.7657625
## se
sqrt(0.0110976413^2+0.0227964189^2)

### overlap weighting

km_ow <- survfit(Surv(as.numeric(surv),death) ~ qsmk, data = nhefs, weights = ow)
km_ow %>% ggsurvfit()
##
##estimate
sum(c(1,-1)*km_ow$surv[c(193,279)])
sqrt(sum(km_ow$std.err[c(193,279)]^2))
##
sum(c(1,-1)*km_ipw$surv[c(193,279)])
sqrt(sum(km_ipw$std.err[c(193,279)]^2))
##
sum(c(1,-1)*km_regular$surv[c(193,279)])
sqrt(sum(km_regular$std.err[c(193,279)]^2))
## IPTW+IPCW
library(pec)
ipcw_out <- ipcw(Surv(as.numeric(surv),1-death) ~ qsmk+age+sex,
                 data = nhefs,
                 method = "nonpar",
                 times = as.numeric(nhefs$surv))
a = 1/(ipcw_out$IPCW.subjectTimes)*ipw
km_ipctw <- survfit(Surv(as.numeric(surv),death) ~ qsmk, data = nhefs, weights = a)
km_ipctw %>% ggsurvfit()                
sum(c(1,-1)*km_ipctw$surv[c(193,279)])
sum(km_ipctw$std.err[c(193,279)]^2)^(0.5)
```

