---
title: "Stat 695 Project Markdown File"
author: "Dylan Clarke"
date: "2025-05-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

We begin by loading in the data and packages we need for this project.

```{r preliminaries, echo=FALSE}
library(causaldata)
library(survival)
library(ggplot2)
library(tidyverse)
library("survminer")
library("splitstackshape")
library(olsrr)
library(ipw)

setwd("~/R/win-library/4.0/Causal Survival")

nhefs = causaldata::nhefs_complete
```

Now we clean the data a bit to get it ready. This creates the survival variable we will use.

```{r cleaning}
start_date = ISOdate(1983,1,1)
nhefs$death_date = ISOdate(paste0(19,nhefs$yrdth),nhefs$modth,nhefs$dadth)
len_study = as.numeric(ISOdate(1993,1,1) - start_date)
nhefs$surv = as.numeric(nhefs$death_date - start_date)
nhefs$surv[is.na(nhefs$surv)] = len_study
nhefs$survival = with(nhefs,Surv(surv,death))

nhefs$survtime <- ifelse(nhefs$death==0, 120, 
                         (nhefs$yrdth-83)*12+nhefs$modth)
```

We also create a second subject-time dataset.

```{r subject time data}
nhefs.surv <- expandRows(nhefs, "survtime", drop=F) 
nhefs.surv$time <- sequence(rle(nhefs.surv$seqn)$lengths)-1
nhefs.surv$event <- ifelse(nhefs.surv$time==nhefs.surv$survtime-1 & 
                             nhefs.surv$death==1, 1, 0)
nhefs.surv$timesq <- nhefs.surv$time^2
```

## Inital Exploration of the Data

## Correlations

```{r corr with qsmk}
cor(Filter(is.numeric,nhefs))

cor(as.numeric(nhefs$sex),nhefs$qsmk)
cor(as.numeric(nhefs$race),nhefs$qsmk)
cor(as.numeric(nhefs$active),nhefs$qsmk)
cor(as.numeric(nhefs$active),nhefs$death)
cor(as.numeric(nhefs$exercise),nhefs$qsmk)
cor(as.numeric(nhefs$exercise),nhefs$death)
```
### Selecting Covariates

```{r select covariates}
sapply(nhefs[,24:57],try(table,silent = TRUE))

qsmk_mod = glm(qsmk~ sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + price82 + tax82 + pepticulcer + alcoholpy + otherpain  + exercise + marital,
               data = nhefs, family = binomial)

death_mod = glm(death~sex + age + race + income + school + ht + wt82 + smokeintensity + smokeyrs + price82 + tax82 + bronch + pepticulcer + hayfever + alcoholpy + alcoholfreq + alcoholtype + alcoholhowmuch + headache + otherpain + nerves + infection + active  + exercise + cholesterol,
               data = nhefs, family = binomial)

ols_step_backward_p(qsmk_mod)
```


##Inverse Probability Weights (IPTW, IPCW)

```{r IPW}
#DOES NOT RUN
#IPTW
ipw_treat = ipwpoint(
  exposure   = qsmk,
  family     = "binomial",
  link       = "logit",
  denominator = ~ sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
  numerator   = ~ 1,
  data       = nhefs
)

qsmk_mod = glm(qsmk~ sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
               data = nhefs, family = binomial)

nhefs$p = fitted(qsmk_mod)

pqsmk = mean(nhefs$qsmk)  
nhefs$iptw <- ifelse(nhefs$qsmk==1,  pqsmk/nhefs$p,  
                                (1-pqsmk)/(1-nhefs$p))

#IPCW
nhefs$censored = 1 - nhefs$death
mod_cens = coxph(
  Surv(surv, death, type = "right") ~ qsmk + sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
  data = nhefs
)

bh = basehaz(mod_cens, centered = FALSE)

lp = predict(mod_cens, type = "lp")

H0_at_t <- approx(
  x = bh$time,
  y = bh$hazard,
  xout = nhefs$surv,
  method = "constant",
  rule = 2
)$y

nhefs$p_cens = exp(- H0_at_t * exp(lp))

pC = mean(nhefs$death == 0)
nhefs$ipcw = ifelse(
  nhefs$death == 0,
  pC / nhefs$p_cens,
  1 / (1 - nhefs$p_cens)
)

nhefs$w = nhefs$ipcw * nhefs$iptw

#Version 2
p.denom <- glm(qsmk ~ sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital, 
               data=nhefs, family=binomial())
nhefs$pd.qsmk <- predict(p.denom, nhefs, type="response")

p.num <- glm(qsmk ~ 1, data=nhefs, family=binomial())
nhefs$pn.qsmk <- predict(p.num, nhefs, type="response")

# computation of estimated weights
nhefs$sw.a2 <- ifelse(nhefs$qsmk==1, nhefs$pn.qsmk/nhefs$pd.qsmk,
                     (1-nhefs$pn.qsmk)/(1-nhefs$pd.qsmk))
summary(nhefs$sw.a)
```

## Propensity Scores

```{r}
qsmk_mod = glm(qsmk~ sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
               data = nhefs, family = binomial)

nhefs$p = fitted(qsmk_mod)
```


## Aalen Model

```{r aalen}
#Baseline Aalen
aalen_model_uw = aareg(survival~ qsmk + sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
      data = nhefs)

aalen_model_uw_2 = aareg(survival~ qsmk + sex + age + wt82_71 + marital,
      data = nhefs)

times = aalen_model_uw$times

saa = summary(aalen_model_uw)
saa2 = summary(aalen_model_uw_2)

coef = aalen_model_uw$coefficient

Lambda0 = coef[,"Intercept"]
Bqmk = coef[, "qsmk"]
BS = coef[,"sex1"]
BR = coef[,"race1"]
BA = coef[, "age"]
BSc = coef[,"school"]
BH = coef[,"ht"]
BW = coef[,"wt82"]
BWc = coef[,"wt82_71"]
BSi = coef[,"smokeintensity"]
BSy = coef[,"smokeyrs"]
BPu = coef[,"pepticulcer"]
BAy = coef[,"alcoholpy"]
BO = coef[,"otherpain"]
BE1 = coef[,"exercise1"]
BE2 = coef[,"exercise2"]
BM = coef[,"marital"]

n = nrow(nhefs)
S0_mat = matrix(0, nrow = n, ncol = length(times))
S1_mat = matrix(0, nrow = n, ncol = length(times))

for (j in seq_along(times)) {
  linpred0 =          BS[j]*(nhefs$sex == 1) +
    BR[j]*(nhefs$race == 1) +
    BA[j]*(nhefs$age == 1) +
    BSc[j]*nhefs$school +
    BH[j]*nhefs$ht +
    BW[j]*nhefs$wt82 +
    BWc[j]*nhefs$wt82_71 + 
    BSi[j]*nhefs$smokeintensity +
    BSy[j]*nhefs$smokeyrs +
    #BP[j]*nhefs$price82 +
    BPu[j]*nhefs$pepticulcer +
    BAy[j]*nhefs$alcoholpy +
    BO[j]*nhefs$otherpain +
    BE1[j]*(nhefs$exercise == 1) +
    BE2[j]*(nhefs$exercise == 2) +
    BM[j]*nhefs$marital
  linpred1 = Bqmk[j] +  BS[j]*(nhefs$sex == 1) +
    BR[j]*(nhefs$race == 1) +
    BA[j]*(nhefs$age == 1) +
    BSc[j]*nhefs$school +
    BH[j]*nhefs$ht +
    BW[j]*nhefs$wt82 +
    BWc[j]*nhefs$wt82_71 + 
    BSi[j]*nhefs$smokeintensity +
    BSy[j]*nhefs$smokeyrs +
    #BP[j]*nhefs$price82 +
    BPu[j]*nhefs$pepticulcer +
    BAy[j]*nhefs$alcoholpy +
    BO[j]*nhefs$otherpain +
    BE1[j]*(nhefs$exercise == 1) +
    BE2[j]*(nhefs$exercise == 2) +
    BM[j]*nhefs$marital
  
  S0_mat[, j] <- exp(-Lambda0[j] - linpred0)
  S1_mat[, j] <- exp(-Lambda0[j] - linpred1)
}

S0 <- colMeans(S0_mat,na.rm = TRUE)
S1 <- colMeans(S1_mat, na.rm = TRUE)

tau = max(times)

rmst0 = sum( (S0[-1] + S0[-length(S0)])/2 * diff(times) )
rmst1 = sum( (S1[-1] + S1[-length(S1)])/2 * diff(times) )

ate_rmst = rmst1 - rmst0

#IPW Aalen
aalen_model_w = aareg(survival~ qsmk + sex + race + age + school + ht + wt82 + wt82_71 + smokeintensity + smokeyrs + pepticulcer + alcoholpy + otherpain  + exercise + marital,
      data = nhefs, weights = sw.a2)

times = aalen_model_w$times

coef = aalen_model_w$coefficient

Lambda0 = coef[,"Intercept"]
Bqmk = coef[, "qsmk"]
BS = coef[,"sex1"]
BR = coef[,"race1"]
BA = coef[, "age"]
BSc = coef[,"school"]
BH = coef[,"ht"]
BW = coef[,"wt82"]
BWc = coef[,"wt82_71"]
BSi = coef[,"smokeintensity"]
BSy = coef[,"smokeyrs"]
BPu = coef[,"pepticulcer"]
BAy = coef[,"alcoholpy"]
BO = coef[,"otherpain"]
BE1 = coef[,"exercise1"]
BE2 = coef[,"exercise2"]
BM = coef[,"marital"]

n = nrow(nhefs)
S0_mat = matrix(0, nrow = n, ncol = length(times))
S1_mat = matrix(0, nrow = n, ncol = length(times))

for (j in seq_along(times)) {
  linpred0 =          BS[j]*(nhefs$sex == 1) +
    BR[j]*(nhefs$race == 1) +
    BA[j]*(nhefs$age == 1) +
    BSc[j]*nhefs$school +
    BH[j]*nhefs$ht +
    BW[j]*nhefs$wt82 +
    BWc[j]*nhefs$wt82_71 + 
    BSi[j]*nhefs$smokeintensity +
    BSy[j]*nhefs$smokeyrs +
    #BP[j]*nhefs$price82 +
    BPu[j]*nhefs$pepticulcer +
    BAy[j]*nhefs$alcoholpy +
    BO[j]*nhefs$otherpain +
    BE1[j]*(nhefs$exercise == 1) +
    BE2[j]*(nhefs$exercise == 2) +
    BM[j]*nhefs$marital
  linpred1 = Bqmk[j] +  BS[j]*(nhefs$sex == 1) +
    BR[j]*(nhefs$race == 1) +
    BA[j]*(nhefs$age == 1) +
    BSc[j]*nhefs$school +
    BH[j]*nhefs$ht +
    BW[j]*nhefs$wt82 +
    BWc[j]*nhefs$wt82_71 + 
    BSi[j]*nhefs$smokeintensity +
    BSy[j]*nhefs$smokeyrs +
    #BP[j]*nhefs$price82 +
    BPu[j]*nhefs$pepticulcer +
    BAy[j]*nhefs$alcoholpy +
    BO[j]*nhefs$otherpain +
    BE1[j]*(nhefs$exercise == 1) +
    BE2[j]*(nhefs$exercise == 2) +
    BM[j]*nhefs$marital
  
  S0_mat[, j] <- exp(-Lambda0[j] - linpred0)
  S1_mat[, j] <- exp(-Lambda0[j] - linpred1)
}

S0 <- colMeans(S0_mat,na.rm = TRUE)
S1 <- colMeans(S1_mat, na.rm = TRUE)

tau = max(times)

rmst0 = sum( (S0[-1] + S0[-length(S0)])/2 * diff(times) )
rmst1 = sum( (S1[-1] + S1[-length(S1)])/2 * diff(times) )

ate_rmst = rmst1 - rmst0
```

## Marginal Structural Models

```{r}
# creation of person-month data
nhefs.ipw = expandRows(nhefs, "survtime", drop=F) 
nhefs.ipw$time = sequence(rle(as.numeric(nhefs.ipw$seqn))$lengths)-1
nhefs.ipw$event = ifelse(nhefs.ipw$time==nhefs.ipw$survtime-1 & 
                            nhefs.ipw$death==1, 1, 0)
nhefs.ipw$timesq <- nhefs.ipw$time^2

# fit of weighted hazards model
ipw.model <- glm(event==0 ~ qsmk + I(qsmk*time) + I(qsmk*timesq) + 
                   time + timesq, family=binomial(), weight=w,
                 data=nhefs.ipw)
summary(ipw.model)

ipw.model_2 <- glm(event==0 ~ qsmk + I(qsmk*time) + I(qsmk*timesq) + 
                   time + timesq, family=binomial(), weight=sw.a,
                 data=nhefs.ipw)
summary(ipw.model)

# creation of survival curves
ipw.qsmk0 <- data.frame(cbind(seq(0, 119),0,(seq(0, 119))^2))
ipw.qsmk1 <- data.frame(cbind(seq(0, 119),1,(seq(0, 119))^2))

colnames(ipw.qsmk0) <- c("time", "qsmk", "timesq")
colnames(ipw.qsmk1) <- c("time", "qsmk", "timesq")

# assignment of estimated (1-hazard) to each person-month */
ipw.qsmk0$p.noevent0 <- predict(ipw.model, ipw.qsmk0, type="response")
ipw.qsmk1$p.noevent1 <- predict(ipw.model, ipw.qsmk1, type="response")

# computation of survival for each person-month
ipw.qsmk0$surv0 <- cumprod(ipw.qsmk0$p.noevent0)
ipw.qsmk1$surv1 <- cumprod(ipw.qsmk1$p.noevent1)

# some data management to plot estimated survival curves
ipw.graph <- merge(ipw.qsmk0, ipw.qsmk1, by=c("time", "timesq"))
ipw.graph$survdiff <- ipw.graph$surv1-ipw.graph$surv0

# plot
ggplot(ipw.graph, aes(x=time, y=surv)) + 
  geom_line(aes(y = surv0, colour = "0")) + 
  geom_line(aes(y = surv1, colour = "1")) + 
  xlab("Months") + 
  scale_x_continuous(limits = c(0, 120), breaks=seq(0,120,12)) +
  scale_y_continuous(limits=c(0.6, 1), breaks=seq(0.6, 1, 0.2)) +
  ylab("Survival") + 
  ggtitle("Survival from IP weighted hazards model") + 
  labs(colour="A:") +
  theme_bw() + 
  theme(legend.position="bottom")

S0 = ipw.qsmk0$surv0 
S1 = ipw.qsmk1$surv1

rmst0_msm = sum( (S0[-1] + S0[-length(S0)])/2 )
rmst1_msm = sum( (S1[-1] + S1[-length(S1)])/2 )

ate_rmst_msm = (rmst1_msm - rmst0_msm) * (3653/120)
```

